---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Observation

The goal of Observation is to provide a free and convenient way of collecting and processing physical activity-specific direct observation data.

## Installation

You can install Observation from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("paulhibbing/Observation")
```

## Example

You can run the program with:

```{r example}
observation_data <- data_collection_program()
```

This will collect direct observation data as you direct. It will also pre-classify the intensity of each activity based on the decision tree described by \url{Hibbing et al.}{https://www.ncbi.nlm.nih.gov/pubmed/29135657}. You can then update the classification by cross-referencing the \url{Compendium of Physical Activities}{https://sites.google.com/site/compendiumofphysicalactivities/} using the following:

```{r example2}
observation_data_complete <- compendium_reference(observation_data)
```
