---
title: "Observation: Collect and Process Physical Activity Direct Observation Data"
author: "Paul R. Hibbing"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Observation: Collect and Process Physical Activity Direct Observation Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1. Overview

The purpose of `Observation` is to provide a free and easy-to-use tool for working with direct observation data in the realm of physical activity. The package has two parts, the first for collecting data, and the second for processing it. Each part has a major function associated with it (i.e., `data_collection_program` to collect data and `compendium_reference` to process). The method is presented in detail in the text and supplemental material of [Hibbing et al. (2018)](https://www.ncbi.nlm.nih.gov/pubmed/29135657). This vignette is designed to provide instructions for use to users who may not be familiar with R.

## 2. Download and Installation

### 2.1 A Critical Note

The following sections provide basic instructions for installing software. For those who already have R and RStudio installed, please note that **having current versions of the applications is critical for the functionality of `Observation`**. If you experience issues with `Observation` (e.g. information is printed in the console, not dialog boxes), one or both applications may be out of date. It is recommended to first update RStudio to see if your issues resolve (go to `Help -> Check for Updates`). If you need to update R, the `installr` package has a useful `updateR` function to guide your installation.

### 2.2 R and RStudio

To use `Observation`, you must naturally have R installed, which you can download for free at https://www.r-project.org/. It is recommended to also download and install RStudio for free from [https://www.rstudio.com](https://www.rstudio.com/products/rstudio/download/).

R is a programming language, whereas RStudio is an "integrated development environment." That is, RStudio includes additional functionality to streamline the way you use R.

### 2.3 The Observation Package

To install the `Observation` package, open Rstudio and type into the console (bottom left pane) `install.packages("Observation")`. Press enter to execute the command, and the package will be automatically installed.

## 3 Using the Observation Package

You can use `Observation` in several ways, but the most recommendable is with a `.R` file. This is simply a text file that stores R commands. To create a `.R` file, open RStudio and select `File -> New File -> R Script`. Now all you need to do is populate the file with code and run it (by highlighting the code and clicking `Run` in the top right of the pane). You can copy the below code into your file to try it out, and then customize the code to meet your needs.

### 3.1 Code

```{r example, eval = FALSE}

# First, you need to "attach" the package. You can think of this as loading it. 
# This step is technically optional, but to use the package functions without
# it, you need to write "Observation::" before each command, e.g.
# "Observation::data_collection_program()"

library(Observation)

# Now it's time to run the Observation program, which will guide you through the
# data collection process described by Hibbing et al. (2018).

# data_collection_program()
  # ^This only runs the program, but does not store the data.
  # You will want to define an object that stores the data you collect.
  # To do so, you provide the name ("my_data") and use the "<-" operator
  # to assign the results of data_collection_program() to an object of
  # that name.

my_data <- data_collection_program()

# You can view your work with

View(my_data)

# There is also a sample data set you can examine with

data(example_data, package = "Observation")
View(example_data)

# The format of "my_data" and "example_data" (and any other data
# collected with data_collection_program()) will be the same. Information
# about what each column represents is available with

help(example_data, package = "Observation")

# Once you are finished collecting data, you should save it to an external file.
# There are a lot of options both for saving in different formats, and for
# managing data from multiple participants. However, this vignette is not
# intended as a tutorial for those types of tasks, and you probably already
# have a system you would rather use at that level. Thus, a minimal example is
# provided here, and the work of determining the appropriate management scheme
# for a given study is left to the reader.

write.csv(my_data, file = "My Example Observation Data.csv", row.names = FALSE)

# Naturally, you should change the filename in the above code to suit your
# needs, and be careful to change the filename each time you run your code, to
# avoid overwriting previously-collected data files. You can easily automate the
# data saving process to avoid hazards, but again, that is beyond the scope of
# this vignette.

# Next, it is time to process the data, again via the scheme described by
# Hibbing et al. (2018), in reference to the Compendium of Physical Activities.
# As before, you need to assign the processed data to an object via "<-",
# which has been named "my_data_processed" below.

my_data_processed <- compendium_reference(my_data)

# You can save this processed data with similar code as given above.

write.csv(my_data_processed, file = "My Example_Processed.csv", row.names = FALSE)

```

### 3.2 Additional Comments

* It is not necessary to run `compendium_reference` immediately after collecting data. If you store the data you collect, you can load it again later (e.g. using `my_data <- read.csv("My Example Observation Data.csv", stringsAsFactors = FALSE)`) and then run `compendium_reference(my_data)`.

* If you will be performing all your analyses in R, it is recommendable to use `save` and `load`, rather than `write.csv` and `read.csv`.

* If you are planning to move forward with csv files, and it is taking too long to read/write the files, you should install the `data.table` package and use the `fread` and `fwrite` commands instead. Those commands work essentially the same as `read.csv` and `write.csv`, but operate much faster.

* When you are specifying filenames, the safest thing to do is write out the whole file path. For example, instead of writing `file = "My Example Observation Data.csv"` inside `write.csv()`, you may wish to write something more like `file = "C:/Users/Me/Documents/My Example Observation Data.csv"`. Otherwise, the command `getwd()` is helpful for finding files if you're not sure where you saved them.

* The general approach for the `compendium_reference` function is to examine the activity descriptions provided during data collection and determine the intensity category (when it was not previously detected) by cross-referencing the [2011 Compendium of Physical Activities](https://sites.google.com/site/compendiumofphysicalactivities/home).
